Setting up Data

### 2020-05-21 src/SetupData.Rmd
DEBUGGED and CLEANED up the setup rmd and MS_v2 scripts

meta.data = read.csv( "~./MLH1repo/data/clean.Meta.Data.txt", sep = "\t", header=TRUE)
#remove all MAD (peromyscus) from meta.data
#meta.data <- meta.data[ !grepl("MAD", meta.data$mouse) , ]

source("~./MLH1repo/src/CommonFunc_MLH1repo.R" )

@@

#total SC -- the one file should be all I need?
Total.SC_DF = read.xlsx2("~./MLH1repo/data/Total.SC_merge.MLH1.xlsx", header=TRUE, 1)

#Curated Bivdata - this excel file is too big to work with read.xlsx2
#Curated_BivData = read.xlsx2("~./MLH1repo/data/Curated_BivData.xlsx", header=TRUE, 1)

@@

@@

#calculate the correlation across duplicates. First issue is to match up and organize the matching 
#for each fileName - make cols, for MLH1-1 and MLH1-2
#remove Xs, for fileName with 
original_DF_duplicates <- original_DF_duplicates[!grepl("X", original_DF_duplicates$X),]
#make a new DF -- with one fileName and the MLH1-1 and MLH1-2
#  (duplicate / dostinct fileName )

@@

`r length(original_DF_duplicates$Random.Name)` rows are marked as duplicate.

@@

@@

@@

#juvi mice ~ 2 weeks old
juvi_mice <- c(
  "30dec14_WSB_m2","30dec14_WSB_m4","30dec14_WSB_m1","30dec14_WSB_m3","30dec14_WSB_m5",
  "30sep14_G_m1","30sep14_G_m2","30sep14_G_m3","30sep14_G_m4","30sep14_G_m5",
  "12Feb17_LEW_m1","12Feb17_LEW_m2", "13sep17_LEW_m2", "13sep17_LEW_m1","13sep17_LEW_m3",

@@

#make juvi.mice DF
juvi.mice_DF <- MLH1_data[(MLH1_data$mouse %in% juvi_mice),]

#remove juvi mice from big MLH1 -- but Then i need to remake the mouse_table
MLH1_data <- MLH1_data[ ! MLH1_data$mouse %in% juvi_mice, ]

@@

old.mice <- c(
  "12oct15_WSB_m1","9oct17_WSB_m1","1mar18_WSB_m2","21dec14_WSB_m2","21dec14_WSB_m3", "1mar18_WSB_m1",
  "14mar15_G_m1","14mar15_G_m2","24mar15_G_m1","19oct15_G_m1",
  "8may17_LEW_m1","18sep17_LEW_m1","8may17_LEW_m2","20dec16_LEW_m2",
  "5jul17_PERC_m1",

@@

I have made a list of the old mice, but from the effect document the older mice might not be that different.

the excluded TOM, had low quality cells, only 4. The exclude AST, had low poor quality cells, there are 3 other mice.  The excluded CZECH males are poor quality (but near the mean). There are 3 other males.

@@

```{r reorder.DF, echo=FALSE}
#this is most important for plotting

#reorder dataframe
MLH1_data <- MLH1_data %>%
  arrange(strain, sex, mouse) %>%
  mutate(Original.Name = factor(Original.Name) )

@@

```{r Date.imaged, echo=FALSE}

#for issues with the total SC metrics, code in date imaged
#use the add_mouse function

#mini <- MLH1_data[1:500,]

count=1
for(i in MLH1_data$Original.Name){
   # print(i)
    templist= strsplit(i, split="_")[[1]]
    c = (templist[1])
    MLH1_data$image.string[count] <-  c
    #mini$image.date[count] <-  as.Date(c, format='%d%b%y')  #"%d%b%y")
    count= count +1
}
#works better if dat is converted out of image
MLH1_data$date.imaged <- as.Date(MLH1_data$image.string,  format='%d%b%y')

#MLH1_data <- add_mouse2(MLH1_data)

#mini.MLH1_datamouse.table <- ddply(MLH1_data, c("mouse"), summarise,
#                        Nmice = length(unique(mouse)),
#                        Ncells  = length(Random.Name),
#                        #most recent date
#                        oldest = min(date.imaged),
#                        newest = max(date.imaged)
#)

# Mouse Data

Make mouse level tables  (re think this name, AP to distinguish BD -- but this isn't really an issue)

```{r all.data.table, echo=FALSE, eval=FALSE}

#make the table with ALL MLH1 obs
FULL_mouse_table_MLH1 <- ddply(MLH1_data, c("mouse"), summarise,

                        Nmice = length(unique(mouse)),
                        Ncells  = length(nMLH1.foci),
                        mean_co = as.numeric(format(round(  mean(nMLH1.foci), 3 ), nsmall=3) ),
                        cV = cv(nMLH1.foci),
                        var = as.numeric(format(round(   var(nMLH1.foci),3), nsmall=3)),
                        sd   = round(sd(nMLH1.foci), 3),
                        se   = round(sd / sqrt(Ncells), 3)
)

#NA cause by CAST female
#this table isn't used much
```

@@

@@

                        oldest.image.date = min(date.imaged),
                        newest.image.date = max(date.imaged)
                        #quality?
)

@@

@@

@@

i) HetC models: mouse averages, only strains with both sexes, no juvi males, only M.musculus. start with **AP_mouse_table**

```{r HetC.DF.4MM, echo=FALSE}

#variance is still chr
Mouse.Table_HQ$var <- as.numeric(Mouse.Table_HQ$var)

DF.HetC.MixedModel <- Mouse.Table_HQ %>% filter(species == "M.musculus")

#remove the extra merged cols
#DF.HetC.MixedModel <- subset(DF.HetC.MixedModel, select= -c(sex.y,sex.x,
#                            subsp.y,subsp.x,
#                    category.y,category.x, strain.x,
#                    strain.y, mouse.1  )  )

#REMOVE STRAINS WITHOUT FEMALES
#DF.HetC.MixedModel <- DF.HetC.MixedModel %>% filter( subsp != "Cast") %>% filter(strain != "AST")

# These are removed in the analysis
#DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$subsp != "Cast",]
#remove musc strains
#DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "AST",]
#DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "TOM",]
#remove PERC, MOLF, CZECH, (no females)
#DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "CZECH",]
#DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "PERC",]
#F.HetC.MixedModel <- droplevels(DF.HetC.MixedModel)#remove levels not used

#Only us the righ age window
#DF.HetC.MixedModel.HQ <- DF.HetC.MixedModel[( (DF.HetC.MixedModel$age.weeks < 25) &  #(DF.HetC.MixedModel$age.days != 13) ) | is.na(DF.HetC.MixedModel$age.weeks), ] #keeps mice without ages, and removes juvinilles

#table(DF.HetC.MixedModel.HQ$strain, DF.HetC.MixedModel.HQ$sex)

#remove blank rows (which are these?)
#DF.HetC.MixedModel <- DF.HetC.MixedModel[!(is.na(DF.HetC.MixedModel$mouse) | #DF.HetC.MixedModel$mouse==""), ]
#151 mice, no rows with nas for values

#main df with variables coded as chr
#DF.HetC.MixedModel.HQ$subsp <- factor(DF.HetC.MixedModel.HQ$subsp, ordered = FALSE)
#DF.HetC.MixedModel.HQ$sex <- factor(DF.HetC.MixedModel.HQ$sex, ordered = FALSE)
#DF.HetC.MixedModel.HQ$strain <- factor(DF.HetC.MixedModel.HQ$strain, ordered = FALSE)
#DF.HetC.MixedModel.HQ$species <- factor(DF.HetC.MixedModel.HQ$species, ordered = FALSE)

#kable(table(DF.HetC.MixedModel.HQ$sex, DF.HetC.MixedModel.HQ$strain))
#only female male reps

```{r male.specific.table,echo=FALSE}

#AP_mouse_table_w.Ages
#use the above table to pull the male mouse levels
mlh1male_mouse.av <-  Mouse.Table_HQ %>% filter(sex== "male")

#add the Rec.group markers
#mouse.avs_4MM$Rec.group <- ifelse(grepl("PWD", mouse.avs_4MM$strain), 1, 
#                            ifelse(grepl("MSM", mouse.avs_4MM$strain), 1,
#                             ifelse(grepl("SKIVE", mouse.avs_4MM$strain), 1, 0)) )

#formate the table for pluggin into models
```

@@

## Number of mice per category

@@

@@

# Male specific BivData set

not sure if SC.Length.male_table is used for anything.

```{r male.sc.length.strain, echo=FALSE}

#these measure ect should be in the male - mouse 
#isolate all musc and male
Curated_BivData_male <- Curated_BivData[ (Curated_BivData$sex == "male") & (Curated_BivData$species == "M.musculus"),]

#make strain average table
SC.Length.male_table <- ddply(Curated_BivData_male, c("strain"), summarise,
                             mean_SC = mean(chromosomeLength),
                              n_SC = length(Obj.ID),

                   mean.SC_0CO = mean(chromosomeLength[which(hand.foci.count == 0)], na.rm = TRUE),
                  n_0CO =  sum((hand.foci.count == 0), na.rm = TRUE ),
                 # CO2 = sum(hand.foci.count == 2, na.rm=TRUE),

                   mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
                    n_1CO = sum((hand.foci.count == 1), na.rm = TRUE),

                   mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
                   n_2CO = sum((hand.foci.count == 2), na.rm = TRUE),

                  mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
                  n_3CO = sum((hand.foci.count == 3), na.rm = TRUE)

# Cell Level Table

@@

# Reduced Bivalent Dataset (Long and Short)

@@

#2. calq the Q4 threshold
#loop for greping the mouse from file name ect
#each Q is a quartille

#mini.now  <- BivData_from_15biv_cells[1:300,]

for(  i in 1:length( unique(BivData_from_15biv_cells$Obj.ID) ) ){

@@

```{r Look.short.biv.data, echo=FALSE, include=FALSE, eval=FALSE}

#showing SC lengths per cell, with lines for the quartille values.

mini_G.15_15biv_FN <- BivData_from_15biv_cells %>% filter(category == "G male")
#ploting short biv data
G.15_15biv_male <- BivData_from_15biv_cells %>% filter(category == "G male")
mini_G.15_15biv_FN <- sample( G.15_15biv_male$fileName, 10)
mini_G.15_15biv  <- BivData_from_15biv_cells[BivData_from_15biv_cells$fileName %in% mini_G.15_15biv_FN, ]

ppll <- ggplot(mini_G.15_15biv, aes(x = fileName, y=chromosomeLength, color=mouse))+geom_jitter()+facet_wrap(~fileName, scales="free")+ theme(legend.position="none")+
  geom_hline(aes(yintercept= cell.Q1), color='coral') + 
  geom_hline(aes(yintercept= cell.Q2), color='red') +
 geom_hline(aes(yintercept= cell.Q3), color='blue') +   ##actually Q3 might not be a real thing...
 geom_hline(aes(yintercept= cell.Q4), color='black')

@@

```{r table.long.biv, echo=FALSE}

# make a table for the Mixed Models -- from the long.BivData
# Calq mouse averages

#add in all the other stats there are in the other tables

@@

#1) Long Biv for 
#remove all the non-used strains!!
Long.biv_mouse_table_HetC <- Long.biv_mouse_table

Long.biv_mouse_table_HetC <- Long.biv_mouse_table_HetC[Long.biv_mouse_table$strain != "AST",]
Long.biv_mouse_table_HetC <- Long.biv_mouse_table_HetC[Curated_BivData$strain != "TOM",]
Long.biv_mouse_table_HetC <- Long.biv_mouse_table_HetC[Long.biv_mouse_table_HetC$strain != "CZECH",]
Long.biv_mouse_table_HetC$strain <- droplevels(Long.biv_mouse_table_HetC$strain)

#remove na's
Long.biv_mouse_table_HetC <- Long.biv_mouse_table_HetC[!(is.na(Long.biv_mouse_table_HetC$mouse) | Long.biv_mouse_table_HetC$mouse==""), ]

@@

#2) Long Biv for male.poly
#just males, just Musculus
Long.biv_mouse_table_male.poly <- Long.biv_mouse_table

Long.biv_mouse_table_male.poly <- Long.biv_mouse_table_male.poly[Long.biv_mouse_table_male.poly$sex == "male",]

#table(droplevels( Long.biv_mouse_table_male.poly$category) ) #for double checking
```

The above code chunk makes 2 mouse level tables; **Long.biv_mouse_table_HetC** and **Long.biv_mouse_table_male.poly**. Run these through the model systems.

# Mouse Level Single BivData Table

@@

The block below creates clean.BivData, which removes ~600 rows if just SC measures
(This might be important for the proportion code)

table(Curated_BivData$hand.foci.count, exclude = NULL)
#there are ~600 rows with SC and no hand foci count (blah)

#remove bivalent entries without hand foci count values
cleanCurated_BivData <-  Curated_BivData  %>%  filter(hand.foci.count  == 0 | hand.foci.count  == 1 |
                                      hand.foci.count  == 2 | hand.foci.count  == 3 | hand.foci.count  == 4 )

                       ncells = length(unique(fileName)),

                       boxs.IDd = max(boxNumber),
                       nboxes_passed = length(boxNumber),

                       Nbiv = length(Obj.ID), #if there are na's below -- it breaks

                      tot.biv = length( !is.na(hand.foci.count) ),

                       mean.SC_length = mean(chromosomeLength, na.rm = TRUE),
                       var.SC_length = var(chromosomeLength, na.rm = TRUE),
                       sd.SC_length   = round(sd(chromosomeLength), 3),
                       se.SC_length   = round(sd.SC_length / sqrt(Nbiv), 3),

                       #sis.co.ten
                       mean.sis.co.ten = mean(SisCoTen, na.rm = TRUE)
)

mouse.level.BivData_table <- add_sex(mouse.level.BivData_table)
mouse.level.BivData_table <- add_strain(mouse.level.BivData_table)
mouse.level.BivData_table <- add_subsp(mouse.level.BivData_table)
mouse.level.BivData_table <- add_category(mouse.level.BivData_table)

mouse.level.BivData_table$Rec.group <- ifelse(grepl("PWD male", mouse.level.BivData_table$category), 1, 
                              ifelse(grepl("MSM male", mouse.level.BivData_table$category), 1,
                                 ifelse(grepl("SKIVE male", mouse.level.BivData_table$category), 1, 0)))
```

@@

```{r mouse.table.Male.Poly, echo=FALSE}

@@

# Melt Position Table

@@

```{r MELT.block1, echo=FALSE}

@@

```{r SHORT.biv_setup, echo=FALSE, eval=TRUE}

#move to Data.setup

#FROM q1
#Q1.long.biv_1CO <-  real.long.bivData %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 1) %>% filter(strain != "MOLF") %>% filter(strain != "CZECH")
#Q1.short.biv_1CO <-  real.short.bivData %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 1)%>% filter(strain != "MOLF") %>% filter(strain != "CZECH")
#from Q2 IFD
#male.long.biv.2CO_IFD <- real.long.bivData %>% filter(sex== "male") %>% filter(species =="M.musculus") %>% filter(hand.foci.count == 2)
##long biv
#male.long.biv <- real.long.bivData %>% filter(sex== "male") %>% filter(species == "M.musculus") 
#table( droplevels( male.long.biv$strain)  )

#FROM q1 
REAL.short_mouse_table <- ddply(real.short.bivData, c("mouse"), summarise,
                        n_obs = length(Obj.ID),
                        ncells = length(unique(fileName)),
                         mean_SC = mean(chromosomeLength),

                      mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
                       mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
                       mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
                       mean.SC_4CO = mean(chromosomeLength[which(hand.foci.count == 4)], na.rm = TRUE),

                      #make sure 1CO  
                       #IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE
                    mean_F1pos_ABS =  mean(Foci1[which(hand.foci.count == 1)]),
                    mean_F1pos_PER =  mean(PER_Foci_1[which(hand.foci.count == 1)]),

                      meanIFD2CO_raw = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
                     meanIFD2CO_norm = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE)

REAL.short_mouse_table <- add_sex(REAL.short_mouse_table)
REAL.short_mouse_table <- add_strain(REAL.short_mouse_table)
REAL.short_mouse_table <- add_subsp(REAL.short_mouse_table)
REAL.short_mouse_table <- add_category(REAL.short_mouse_table)

REAL.short_mouse_table$sex <- factor(REAL.short_mouse_table$sex, ordered = FALSE)
REAL.short_mouse_table$strain <- factor(REAL.short_mouse_table$strain, ordered = FALSE)
REAL.short_mouse_table$subsp <- factor(REAL.short_mouse_table$subsp, ordered = FALSE)

#divide musc males into groups
REAL.short_mouse_table$Rec.group <- ifelse(grepl("PWD male", REAL.short_mouse_table$category), 1, 
                                  ifelse(grepl("MSM male", REAL.short_mouse_table$category), 1,
                                  ifelse(grepl("SKIVE male", REAL.short_mouse_table$category), 1,
                                           0)))

###LONG BIV

REAL.long_mouse_table <- ddply(real.long.bivData, c("mouse"), summarise,
                       n_obs = length(Obj.ID),
                        ncells = length(unique(fileName)),
                         mean_SC = mean(chromosomeLength),

                      mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
                       mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
                       mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
                       mean.SC_4CO = mean(chromosomeLength[which(hand.foci.count == 4)], na.rm = TRUE),

                      #make sure 1CO,  #IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE
                    mean_F1pos_ABS =  mean(Foci1[which(hand.foci.count == 1)]),
                    mean_F1pos_PER =  mean(PER_Foci_1[which(hand.foci.count == 1)]),

                      meanIFD2CO_raw = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
                     meanIFD2CO_norm = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE)
                         )

REAL.long_mouse_table <- add_sex(REAL.long_mouse_table)
REAL.long_mouse_table <- add_strain(REAL.long_mouse_table)
REAL.long_mouse_table <- add_subsp(REAL.long_mouse_table)
REAL.long_mouse_table <- add_category(REAL.long_mouse_table)

REAL.long_mouse_table$sex <- factor(REAL.long_mouse_table$sex, ordered = FALSE)
REAL.long_mouse_table$strain <- factor(REAL.long_mouse_table$strain, ordered = FALSE)
REAL.long_mouse_table$subsp <- factor(REAL.long_mouse_table$subsp, ordered = FALSE)

#divide musc males into groups
REAL.long_mouse_table$Rec.group <- ifelse(grepl("PWD male", REAL.long_mouse_table$category), 1, 
                                  ifelse(grepl("MSM male", REAL.long_mouse_table$category), 1,
                                  ifelse(grepl("SKIVE male", REAL.long_mouse_table$category), 1,
                                           0)))

#find which one

>>>>

meta.data = read.csv( "~./MLH1repo/data/clean.Meta.Data.txt", sep = "\t", header=TRUE)

@@

#total SC 
Total.SC_DF = read.xlsx2("~./MLH1repo/data/Total.SC_DF.xlsx", header=TRUE, 1)

@@

source("~./MLH1repo/src/CommonFunc_MLH1repo.R" )
```

@@

#calculate the correlation across duplicates. First issue is to match up and organize the matching 
#for each fileName - make cols, for MLH1-1 and MLH1-2
#remove Xs, for fileName with 
original_DF_duplicates <- original_DF_duplicates[!grepl("X", original_DF_duplicates$X),]
#these are images counted across mulitple batches

@@

`r length(original_DF_duplicates$Random.Name)` rows are marked as duplicate.  `r length(original_DF_duplicates$fileName)` rows (images) have been counted across mulitple batches. These are removed in the MLH1_data DF.

@@

`r length(original_DF$fileName[ grepl("X", original_DF$X)] )` images (rows) have X's, they are removed from the main dataset.

@@

# Mouse Data MLH1

@@

#juvi mice ~ 2 weeks old
juvi_mice.list <- c(
  "30dec14_WSB_m2","30dec14_WSB_m4","30dec14_WSB_m1","30dec14_WSB_m3","30dec14_WSB_m5",
  "30sep14_G_m1","30sep14_G_m2","30sep14_G_m3","30sep14_G_m4","30sep14_G_m5",
  "12Feb17_LEW_m1","12Feb17_LEW_m2", "13sep17_LEW_m2", "13sep17_LEW_m1","13sep17_LEW_m3",

@@

#make juvi.mice DF
juvi.mice_DF <- MLH1_data[(MLH1_data$mouse %in% juvi_mice.list),]

#remove juvi mice from big MLH1 -- but Then i need to remake the mouse_table
MLH1_data <- MLH1_data[ ! MLH1_data$mouse %in% juvi_mice.list, ]

@@

old.mice.list <- c(
  "12oct15_WSB_m1","9oct17_WSB_m1","1mar18_WSB_m2","21dec14_WSB_m2","21dec14_WSB_m3", "1mar18_WSB_m1",
  "14mar15_G_m1","14mar15_G_m2","24mar15_G_m1","19oct15_G_m1",
  "8may17_LEW_m1","18sep17_LEW_m1","8may17_LEW_m2","20dec16_LEW_m2",
  "5jul17_PERC_m1",

@@

`r length(old.mice.list)` old mice (> 25 weeks) were dissected, some of the RIKEN musc strains measures only have old mouse observations.

`r length(juvi_mice.list)` juvinile mice were dissected, mostly only for DMC1 counts. Few had good MLH1 cells so they were removed from the main MLH1 DF.

@@

```{r reorder.DF, echo=FALSE}
#this is most important for plotting
#reorder dataframe
MLH1_data <- MLH1_data %>%
  arrange(strain, sex, mouse) %>%
  mutate(Original.Name = factor(Original.Name) )

@@

## MLH1 mouse level tables**

@@

#add date image
count=1
for(i in MLH1_data$Original.Name){
   # print(i)
    templist= strsplit(i, split="_")[[1]]
    c = (templist[1])
    MLH1_data$image.string[count] <-  c
    #mini$image.date[count] <-  as.Date(c, format='%d%b%y')  #"%d%b%y")
    count= count +1
}
#works better if dat is converted out of image
MLH1_data$date.imaged <- as.Date(MLH1_data$image.string,  format='%d%b%y')

@@

                        #dates Imaged not in the DF
                        oldest.image.date = min(date.imaged),
                        newest.image.date = max(date.imaged)
                        #quality?
)

@@

```{r merge.meta_MLH1, echo=FALSE}

#merge - MLH1 with metdata

@@

Set up the dataframes in results or MS.

@@

i) HetC models: mouse averages, only strains with both sexes, no juvi males, only M.musculus.

@@

### Number of mice per category

@@

table(Curated_BivData$hand.foci.count, exclude = NULL)
#there are ~600 rows with SC and no hand foci count (blah)
#Remove bivalent entries without hand foci count values
cleanCurated_BivData <-  Curated_BivData  %>%  filter(hand.foci.count  == 0 | hand.foci.count  == 1 |
                                      hand.foci.count  == 2 | hand.foci.count  == 3 | hand.foci.count  == 4 )

length( (Curated_BivData  %>%  filter(  is.na(hand.foci.count) == T ) )$fileName )

table(cleanCurated_BivData$hand.foci.count, exclude = NULL)

                      Nmice = length(unique(mouse)),
                       Ncells = length(unique(fileName)),
                       boxs.IDd = max(boxNumber),
                       nboxes_passed = length(boxNumber),
                      nbiv = length(Obj.ID),

                      tot.biv = length( !is.na(hand.foci.count) ),
                      # mean.SC_length = mean(chromosomeLength, na.rm = TRUE),
                       mean_SC = as.numeric(format(round(mean(chromosomeLength), 3), nsmall=3) ),
                       var.SC_length = var(chromosomeLength, na.rm = TRUE),
                       sd.SC_length   = round(sd(chromosomeLength), 3),
                       se.SC_length   = round(sd.SC_length / sqrt(nbiv), 3),
                        SC.cV = cv(chromosomeLength),

                  proportion_0CO = (n0CO / nbiv),
                  proportion_1CO = (n1CO / nbiv),
                 proportion_2CO = (n2CO / nbiv),
                     proportion_3CO = (n3CO / nbiv),

                #mean 1CO.pos
                mean_1CO.pos = mean(Foci1[which(hand.foci.count == 1)], na.rm = TRUE),
                mean_1CO.pos.PER = mean(PER_Foci_1[which(hand.foci.count == 1)], na.rm = TRUE),
                #CO position things, requires the melt?
              norm.first.foci.pos =  mean(PER_Foci_1[which(hand.foci.count == 1)]),

                mean_IFD.2CO_ABS = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
                 mean_IFD.2CO_PER = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE),
                # IFD things
             mean_IFD_ABS = as.numeric(format(round(mean(IFD1_ABS, na.rm =TRUE), 3), nsmall=3)),
             mean_IFD_PER = as.numeric(format(round(mean(IFD1_PER, na.rm = TRUE), 3), nsmall=3)),

              #sis.co.ten
              mean.siscoten = mean(SisCoTen, na.rm=TRUE),
              mean.siscoten_1CO = mean(SisCoTen[which(hand.foci.count == 1)], na.rm = TRUE),
              mean.siscoten_2CO = mean(SisCoTen[which(hand.foci.count == 2)], na.rm = TRUE),
              mean.siscoten_3CO = mean(SisCoTen[which(hand.foci.count == 3)], na.rm = TRUE),

              mean.telo.dist = mean(telo_dist, na.rm=TRUE),
              mean.cent.dist = mean(dis.cent, na.rm=TRUE),
                 #sis.co.ten
                       mean.sis.co.ten = mean(SisCoTen, na.rm = TRUE)
)

mouse.level.BivData_table <- add_sex(mouse.level.BivData_table)
mouse.level.BivData_table <- add_strain(mouse.level.BivData_table)
mouse.level.BivData_table <- add_subsp(mouse.level.BivData_table)
mouse.level.BivData_table <- add_species(mouse.level.BivData_table)

mouse.level.BivData_table <- add_category(mouse.level.BivData_table)

mouse.level.BivData_table$Rec.group <- ifelse(grepl("PWD male", mouse.level.BivData_table$category), 1, 
                              ifelse(grepl("MSM male", mouse.level.BivData_table$category), 1,
                                 ifelse(grepl("SKIVE male", mouse.level.BivData_table$category), 1, 0)))

```{r CurBiv.Mouse.av.TABLE, echo=FALSE, eval=FALSE}

#18jul19_31may19_SKIVE_m2_sp1_13.2_rev_8 is missing the 1CO position -- remove for now
cleanCurated_BivData <- cleanCurated_BivData %>% filter(fileName != "18jul19_31may19_SKIVE_m2_sp1_13.2_rev_8")

#cleanCurated_BivData
#removed the no hand foci count

#there's alread a Mouse.Table_BivData

mouse.BivData_avs_4MM <- ddply(cleanCurated_BivData, c("mouse"), summarise,
                        Nmice = length(unique(mouse)),
                        Ncells  = length(unique( fileName) ),
                        nbiv = length(Obj.ID),

                        mean_SC = as.numeric(format(round(mean(chromosomeLength), 3), nsmall=3) ),
                        cV = cv(chromosomeLength),
                        var = as.numeric(format(round(   var(chromosomeLength),3), nsmall=3) ),
                        sd   = round(sd(chromosomeLength), 3),
                        se   = round(sd / sqrt(Ncells), 3),

              #can't figure out how to get this format to work  
                       n0CO = sum(length( which(hand.foci.count == 0) )  ),
                       n1CO = sum(length( which(hand.foci.count == 1) )  ),
                      n2CO = sum(length( which(hand.foci.count == 2) )  ),
                      n3CO = sum(length( which(hand.foci.count == 3) )  ),
              #     
                  proportion_0CO = (n0CO / nbiv),
                  proportion_1CO = (n1CO / nbiv),
                 proportion_2CO = (n2CO / nbiv),
                     proportion_3CO = (n3CO / nbiv),

                #mean 1CO.pos
                mean_1CO.pos = mean(Foci1[which(hand.foci.count == 1)], na.rm = TRUE),
                mean_1CO.pos.PER = mean(PER_Foci_1[which(hand.foci.count == 1)], na.rm = TRUE),
                #CO position things, requires the melt?
              norm.first.foci.pos =  mean(PER_Foci_1[which(hand.foci.count == 1)]),

                # IFD things
            #  mean_IFD_ABS = as.numeric(format(round(mean(IFD1_ABS, na.rm =TRUE), 3), nsmall=3)),
            # mean_IFD_PER = as.numeric(format(round(mean(IFD1_PER, na.rm = TRUE), 3), nsmall=3)),

              #sis.co.ten!!
              mean.siscoten = mean(SisCoTen, na.rm=TRUE),
              mean.siscoten_1CO = mean(SisCoTen[which(hand.foci.count == 1)], na.rm = TRUE),
              mean.siscoten_2CO = mean(SisCoTen[which(hand.foci.count == 2)], na.rm = TRUE),
              mean.siscoten_3CO = mean(SisCoTen[which(hand.foci.count == 3)], na.rm = TRUE),

#vari
mouse.avs_4MM <- add_sex(mouse.avs_4MM)
mouse.avs_4MM <- add_strain(mouse.avs_4MM)
mouse.avs_4MM <- add_subsp(mouse.avs_4MM)
mouse.avs_4MM <- add_category(mouse.avs_4MM)
mouse.avs_4MM <- add_species(mouse.avs_4MM)

#unordered for the mixed models (move this to the MM blocks...?)
#keep order for the plots?
mouse.avs_4MM$subsp <- factor(mouse.avs_4MM$subsp, ordered = FALSE)
mouse.avs_4MM$sex <- factor(mouse.avs_4MM$sex, ordered = FALSE)
mouse.avs_4MM$strain <- factor(mouse.avs_4MM$strain, ordered = FALSE)

#REMOVE JUVI mice
juvi.mouse_CurBiv_table <- mouse.avs_4MM[mouse.avs_4MM$mouse  %in% juvi_mice.list, ]  
mouse.avs_4MM <- mouse.avs_4MM[! mouse.avs_4MM$mouse  %in% juvi_mice.list, ]

#assign high and low
#divide musc males into groups
mouse.avs_4MM$Rec.group <- ifelse(grepl("PWD male", mouse.avs_4MM$category), 1, 
                                ifelse(grepl("MSM male", mouse.avs_4MM$category), 1,
                                 ifelse(grepl("SKIVE male", mouse.avs_4MM$category), 1, 0)))

mouse.BivData.Table <- mouse.avs_4MM
```

@@

## Cell Level Table BivData

@@

## Reduced Bivalent Dataset (Long and Short)

@@

#2. calq the Q4 threshold
#loop for greping the mouse from file name ect
#each Q is a quartille
for(  i in 1:length( unique(BivData_from_15biv_cells$Obj.ID) ) ){

@@

The above chunk makes the *real.short.bivData* and *real.long.BivData* BivData.DFs.

```{r plot.short.biv.data, echo=FALSE, include=FALSE, eval=FALSE}

#showing SC lengths per cell, with lines for the quartille values.
ggplot(mini_G.15_15biv, aes(x = fileName, y=chromosomeLength, color=mouse))+geom_jitter()+facet_wrap(~fileName, scales="free")+ theme(legend.position="none")+
  geom_hline(aes(yintercept= cell.Q1), color='coral') + 
  geom_hline(aes(yintercept= cell.Q2), color='red') +
 geom_hline(aes(yintercept= cell.Q3), color='blue') +   ##actually Q3 might not be a real thing...
 geom_hline(aes(yintercept= cell.Q4), color='black')

@@

The chunck below makes the mouse level tabeles for the 2 reduced DFs.

```{r Mouse.Long.Biv_table, echo=FALSE}
# make a table for the Mixed Models -- from the long.BivData

@@

#1) Long Biv for

#remove all the non-used strains!!
#Long.biv_mouse_table_HetC <- Long.biv_mouse_table
#Long.biv_mouse_table_HetC <- Long.biv_mouse_table_HetC[Long.biv_mouse_table$strain != "AST",]
#Long.biv_mouse_table_HetC <- Long.biv_mouse_table_HetC[Curated_BivData$strain != "TOM",]
#Long.biv_mouse_table_HetC <- Long.biv_mouse_table_HetC[Long.biv_mouse_table_HetC$strain != "CZECH",]
#Long.biv_mouse_table_HetC$strain <- droplevels(Long.biv_mouse_table_HetC$strain)

#remove na's
#Long.biv_mouse_table_HetC <- #Long.biv_mouse_table_HetC[!(is.na(Long.biv_mouse_table_HetC$mouse) | #Long.biv_mouse_table_HetC$mouse==""), ]

@@

#2) Long Biv for male.poly
#just males, just Musculus
#Long.biv_mouse_table_male.poly <- Long.biv_mouse_table
#Long.biv_mouse_table_male.poly <- #Long.biv_mouse_table_male.poly[Long.biv_mouse_table_male.poly$sex == "male",]
#table(droplevels( Long.biv_mouse_table_male.poly$category) ) #for double checking
```

```{r Mouse.Short.Biv_table, echo=FALSE}
# make a table for the Mixed Models -- from the long.BivData

Short.biv_mouse_table <- ddply(real.short.bivData, c("mouse"), summarise,

Short.biv_mouse_table <- add_sex(Short.biv_mouse_table)
Short.biv_mouse_table <- add_strain(Short.biv_mouse_table)
Short.biv_mouse_table <- add_subsp(Short.biv_mouse_table)
Short.biv_mouse_table <- add_category(Short.biv_mouse_table)

#unordered for the mixed models (move this to the MM blocks...?)
Short.biv_mouse_table$subsp <- factor(Short.biv_mouse_table$subsp, ordered = FALSE)
Short.biv_mouse_table$sex <- factor(Short.biv_mouse_table$sex, ordered = FALSE)
Short.biv_mouse_table$strain <- factor(Short.biv_mouse_table$strain, ordered = FALSE)

#rec group
Short.biv_mouse_table$Rec.group <- ifelse(grepl("PWD", Short.biv_mouse_table$category), 1, 
                            ifelse(grepl("MSM", Short.biv_mouse_table$category), 1,
                             ifelse(grepl("SKIVE", Short.biv_mouse_table$category), 1, 0)) )

The above 2 plots make mouse level tables for the reduced single biv datasets.

## Mouse Level Single BivData Table

@@

The block below creates clean.BivData, which removes ~600 rows with just SC measures and no position data.

@@

```{r mouse.table.Male.Poly, echo=FALSE, eval=FALSE}

#this makes a redundant table

@@

## BivData Melt Position Table

@@

```{r BivData.MELT.block1, echo=FALSE}

@@

The above chunk makes the MELT table (*MELT_Curated_BivData*), which can be used for Foci position comparison -- But I might not really use this for that.
